---
title: "01-many-sheets"
author: "Jameel Alsalam"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("readxl")
```

## Read data with readxl

Leaning on a readxl [vignette](http://readxl.tidyverse.org/articles/articles/readxl-workflows.html)

Using a workbook that I adapted from an example workbook included in the readxl package in variations.

```{r readxl_examples}
# list sheets
excel_sheets("sheets/deaths.xlsx")
```

If the data is rectangular and located in the upper-left corner, then things are so easy.

```{r}
read_excel("sheets/deaths.xlsx", sheet = "clean")
```


But often spreadsheets have notes and whitespace surrounding the data.

```{r}
read_excel("sheets/deaths.xlsx", "original")
```

If you know where the rectangle is, its no problem.

```{r}
orig_data <- read_excel("sheets/deaths.xlsx", "original", skip = 4)

orig_data
```


```{r}
nrow(orig_data)
```



## Many sheets (or workbooks)

Reading in many sheets or workbooks at once requires:
1) Using a consistent structure (and a function to read it using `readxl`)
2) Applying that function to each workbook with `purrr`

```{r template}

read_template <- function(wb, sheet) {
  # read metadata
  metadata <- read_excel(wb, sheet, range = "A1:B3", 
                         col_names = c("meta", "value")) %>% 
    spread(meta, value) %>%
    as.list()
  
  # read data and append metadata
  data <- read_excel(wb, sheet, skip = 4) %>%
    mutate(Category = metadata[["Category"]],
           File = metadata[["File"]],
           Source = metadata[["Source"]])
  
}

arts <- read_template("sheets/deaths.xlsx", "arts")

```

And now this can be applied to multiple sheets with `purrr::map_dfr`

```{r}

deaths <- map_dfr(c("arts", "other"),
                  ~read_template("sheets/deaths.xlsx", .x))

View(deaths)
deaths
```

Note that both 'Categories' are included.



